Type theories as quotient inductive-recursive types

### 2025-02-21 tex/main.tex
Tidy up the use of brackets

The structural rules for type substitution are stipulated by following equality constructors:
\begin{alignat*}{9}
  [\idS]_{\Ty} &: \implicit{\Gamma, A} \; && \sub{A}{\idS} && =^{\Ty\,\Gamma} A
  \qquad\qquad
  [;]_{\Ty}  : \implicit{\Gamma, \Delta, \Theta, \sigma, \tau, A}\; \sub{A}{\sigma ; \tau} =^{\Ty\,\Gamma} \sub{(\sub{A}{\tau})}{\sigma}
  \\
  []\UU &: \implicit{\Gamma, \Delta, \sigma} \; && \sub{\UU}{\sigma} && =^{\Ty\,\Gamma} \UU
\end{alignat*}
%
%\begin{alignat*}{3}
%  [\idS]_{\Ty} & : \implicit{\Gamma, A}                               && \sub{A}{\idS}           && =^{\Ty\,\Gamma}\;A \\
%  [;]_{\Ty}    & : \implicit{\Gamma, \Delta, \Theta, \sigma, \tau, A}\;&& \sub{A}{\sigma ; \tau}  && =^{\Ty\,\Gamma}\; \sub{(\sub{A}{\tau})}{\sigma} \\
%  []\UU      & : \implicit{\Gamma, \Delta, \sigma}                    && \sub{\UU}{\sigma}        && =^{\Ty\,\Gamma}\;\UU
%\end{alignat*}

@@

However, we still have a little bit more work to do before we can declare victory: as we are defining a function on a quotient inductive type, we have to prove that $\subTy{\blank}{\blank}$ is coherent, i.e., that it respects all equality constructors.
%\LT[noinline]{Should we mention any general condition for coherence?}
%\Fredrik[noinline]{Is there one?}
%\LT[noinline]{I think Cubical Agda has sort of conditions? (Cubical Agda does not support rewriting)}
\begin{proposition}\label{prop:coherence-1}
  For every equality constructor $\cc : \sigma = \tau$, we have $\subTy{A}{\sigma} = \subTy{A}{\tau}$.
  \danger
\end{proposition}
\begin{proof}
  The equality constructors $[\idS]_T$, $[;]_T$, and $[]\UU$ turn into definitional equalities for type substitution, so it remains to show that the following identities
  \begin{align*}
    \subTy{A}{\sigma;\idS}          & = \subTy{A}{\sigma},
                                    & \subTy{A}{\idS;\sigma}          & = \subTy{A}{\sigma},
                                     & \subTy{A}{(\sigma;\tau);\gamma} & = \subTy{A}{\sigma; (\tau; \gamma)}, \\
    \subTy{A}{\pi_1(\sigma, t)}      & = \subTy{A}{\sigma}, \\
    \subTy{A}{\sigma; (\tau, t)}    & = \subTy{A}{(\sigma;\tau), \sub{t}{\sigma}},
                                     & \subTy{A}{\sigma}               & = \subTy{A}{\emptyctx},
                                     & \subTy{A}{\sigma}               & = \subTy{A}{\pi_1\sigma, \pi_2\sigma}
  \end{align*}
  hold propositionally.
  The first two identities of the first row hold strictly by \eqref{eq:def-type-subst-1} and \eqref{eq:def-type-subst-2}, while both sides of the third reduce to $\subTy{(\subTy{(\subTy{A}{\gamma})}{\tau})}{\sigma}$ by \eqref{eq:def-type-subst-2}.
  %
  The second row is clause \eqref{eq:def-type-subst-3} of the definition of $\subTy{\blank}{\blank}$.
  %
  We show the remaining cases by induction on $A$.
  However, $\UU$ is the only constructor for $\Ty\;\Gamma$ and each of the identities with $A \judgeq \UU$ reduce to $\UU$.
\end{proof}
\begin{remark}\label{re:coherence-proof}
  In the above proof the first four identities hold strictly by \eqref{eq:def-type-subst-1}--\eqref{eq:def-type-subst-4}, while the last three follow from \eqref{eq:def-type-subst-5} by induction on $A$.
  As we add other type formers, the first four identities will always hold, but the last three will depend on how type substitution is defined.
\end{remark}

@@

\begin{minipage}{.6\textwidth}
\begin{alignat*}{5}
  %\Tel               & :                   && \Ctx \to \Set                                     \\
  \emptytel          & : \implicit{\Gamma}\; && \Tel\;\Gamma \\
  \blank,\blank      & : \implicit{\Gamma}\; && (\Xi : \Tel\;\Gamma) (A : \Ty\;(\Gamma \dplus \Xi)) \to \Tel\;\Gamma
\end{alignat*}
\end{minipage}
\begin{minipage}{.4\textwidth}
\begin{alignat*}{5}
  %\blank & \dplus\blank     && :       (\Gamma : \Ctx) (\Xi : \Tel\;\Gamma) \to \Ctx \\
  \Gamma & \dplus \emptytel && \reduce \Gamma \\
  \Gamma & \dplus (\Xi, A)  && \reduce (\Gamma \dplus \Xi) , A
\end{alignat*}
\end{minipage}
\vskip\belowdisplayskip\noindent
Equipped with telescopes, we can now generalise single lifting with type \eqref{eq:type-of-lifting} to multiple lifting with type
\[
\blank\upuparrows \blank : (\sigma : \Sub\;\Gamma\;\Delta) \to (\Xi : \Tel\;\Delta)
\to \Sub\;(\Gamma \dplus (\subTel{\Xi}{\sigma}))\;(\Delta \dplus \Xi)
\]
which asks for telescope substitution $\subTel{\blank}{\blank} :  \Tel\;\Delta \to \Sub\;\Gamma\;\Delta \to \Tel\;\Gamma$ defined mutually:

@@

\begin{minipage}{.5\textwidth}
\begin{alignat*}{3}
  \subTel{\Xi}{\idS}          & \reduce \Xi \\
  \subTel{\Xi}{\sigma ; \tau} & \reduce \subTel{(\subTel{\Xi}{\tau})}{\sigma}
\end{alignat*}
\end{minipage}
\begin{minipage}{.5\textwidth}
\begin{alignat*}{5}
  & \idS          && \upuparrows \Xi && \reduce \idS \\
  & (\sigma;\tau) && \upuparrows \Xi && \reduce (\sigma \upuparrows \subTel{\Xi}{\tau}) ; (\tau \upuparrows \Xi)
\end{alignat*}
\end{minipage}\vskip\belowdisplayskip\noindent
%whose confluence property is easy to see.

@@

Finally, we note that for $t : \Tm\,\Gamma\,(\Pi\,A\,B)$ and $u : \Tm\,\Gamma\,A$, also `ordinary' application
$t \mathop{\$} u \defeq \subTm{(\mathsf{app}\,t)}{(\idS , t)}$ of type $\Tm\,\Gamma\,(\sub{B}{(\idS , t)})$ can be defined without any transport. This is in contrast to the QIIT situation, where ordinary application is defined as $t \mathop{\$} u \defeq \subTm{(\mathsf{app}\,t)}{(\idS, \alert{\transfib{\Tm\;\Gamma}{[\idS]_{\Tm}^{-1}}{\color{black}t}})}$.
%
%\begin{alignat*}{3}
%  \left<\blank\right> & : \implicit{\Gamma, A} \Tm\;\Gamma\;A \to \Sub\;\Gamma\;(\Gamma, A) \\
%  \left< t \right>    & \defeq (\idS , t),
%\end{alignat*}
%used to define the ordinal application: $t \mathop{\$} u \defeq [ \left< u \right> ]_{\Tm} (\mathsf{app}\,t)$, 
%can be introduced without any transport, which is $(\idS, \alert{\transfib{\Tm\;\Gamma}{[id]_T^{-1}}{\color{black}t}})$ in its QII definition.
Similarly, the structural rule for substitution on $\mathsf{app}$
\[
  \sub{(\mathsf{app}\;t)}{\sigma \uparrow A} \overset{\Pi\beta^{-1}}{=}
  \mathsf{app}\,\lambda(\sub{(\mathsf{app}\;t)}{\sigma \uparrow A}) \overset{[]\lambda^{-1}}{=}
  \mathsf{app}\,\left(\sub{(\lambda\,(\mathsf{app}\,t))}{\sigma}\right) \overset{\Pi\eta^{-1}}{=}
  \mathsf{app}\,(\sub{t}{\sigma})
\]
%\begin{align*}
%  \sub{(\mathsf{app}\;t)}{\sigma \uparrow A} & = \mathsf{app}\,\lambda(\sub{(\mathsf{app}\;t)}{\sigma \uparrow A}) && \text{by $\Pi\beta^{-1}$} \\
%                                              & = \mathsf{app}\,\left(\sub{(\lambda\,(\mathsf{app}\,t))}{\sigma}\right) && \text{by $[]\lambda^{-1}$} \\
%                                              & = \mathsf{app}\,(\sub{t}{\sigma}) && \text{by $\Pi\eta^{-1}$.}
%\end{align*}
can be derived without transport, since \eqref{eq:def-type-subst-8} is now strict.

@@

\begin{remark}
  To state the methods, we cannot avoid using transports in identities that would need transports in its QIIT definition of a type theory --- for example, we have
  \[
    [\idS]_{\Tm}^M : \implicit{\Gamma, \Gamma^M, t, t^M} \;\subTmM{t^M}{\idS^M} =^{\lam{A^M}\Tm^M\;\Gamma^M\;A^M\;t}_{\alert{[\idS]^M_{\Ty}}} \; t^M
  \]
  since identities occurring in a method %(e.g.\ $[\idS]_{\Ty}^M$)
  are propositional just like equality constructors in its QIIT definition.
  However, in the elimination principle for the QIIT definition, the situation is even worse, as we in general need yet another layer of transports for the methods to typecheck over the QIIT constructors.

@@

The elimination principle computes by recursively interpreting each constructor using the corresponding method. For example, we have
\begin{alignat*}{5}
%  \elim_{\Ctx}\;\emptyctx   & \reduce \emptyctx_{\Ctx}                                    && : \Ctx^M\;\emptyctx  \\
  \elim_{\Ctx}\;(\Gamma, A) & \reduce (\elim_{\Ctx}\,\Gamma),_{\Ctx}^{M} (\elim_{\Ty}\,A) . && %\\
%  \elim_{\Ty}\;\UU          & \reduce \UU^M && :\Ty^M\;\Gamma^M\;\UU \\
%  \elim_{\Sub}\;\emptysub   & \reduce \emptysub^M_{\Sub} &&: \Sub^M\;\Gamma^M\;\emptyctx^M_{\Ctx}\;\emptyctx \\
%                            & \vdots
\end{alignat*}
%
Interestingly, a key difference compared to its QIIT counterpart is that $\subTy{\blank}{\blank}$, $\subTm{\blank}{\blank}$, and $\blank\uparrow\blank$ are functions rather than (formed by) constructors, meaning they cannot be eliminated inductively but instead become propositional identities.
%type substitution $[\blank]\blank$ was a constructor and $[\sigma]\,A$ was eliminated to $[\elim_{\Sub}\,\sigma]^M_{\Ty}\,(\elim_{\Ty}\,A)$; similarly for term substitution.
\begin{restatable}{proposition}{propProvableFunc}\label{prop:provable-function-clauses}
  Given the motives and the methods, we can prove
\begin{alignat*}{5}
  \elim_{\Ty}[] & : \elim_{\Ty}\,(\subTy{A}{\sigma}) && = && \subTyM{(\elim_{\Ty}\,A)}{\elim_{\Sub}\,\sigma} \\
  \elim_{\Tm}[] & : \elim_{\Tm}\,(\subTm{t}{\sigma}) && =^{\lam{A^M}\Tm^M\;\Gamma^M\;A^M\;t}_{\alert{\elim_{\Ty}[]}} && \subTmM{(\elim_{\Tm}\,t)}{\elim_{\Sub}\,\sigma} \\
  \elim_{\Sub}[] & : \elim_{\Sub}(\sigma\uparrow A) && =^{\lam{B^M}\Sub^M\;(\Gamma^M,^M B^M)\;(\Delta^M, A^M)\;(\sigma \uparrow A)}_{\alert{\elim_{\Ty}[]}} && (\elim_{\Sub} \sigma) \uparrow^M (\elim_{\Ty}\,A)
\end{alignat*}
\end{restatable}

@@

\begin{alignat*}{5}
  ;\text{-}\textsf{idr}^M   & : \sigma^M ;^M \idS^M && =^{\Sub^M\,\Gamma^M\,\Delta^M}_{\alert{;\text{-}\mathsf{idr}}} && \sigma^M \\
  ;\text{-}\textsf{idl}^M   & : \idS^M ;^M \sigma^M && =^{\Sub^M\,\Gamma^M\,\Delta^M}_{\alert{;\text{-}\mathsf{idl}}} && \sigma^M \\
  ;\text{-}\textsf{assoc}^M & : (\sigma^M ;^M \tau^M) ;^M \gamma^M   && =^{\Sub^M\,\Gamma^M\,\Delta^M}_{\alert{;\text{-}\mathsf{assoc}}} && \sigma^M ;^M (\tau^M ;^M \gamma^M) \\
  \mathsf{concat}^M         & : \sigma^M ;^M (\tau^M ,_{\Sub}^M t^M) && =^{\Sub^M\,\Gamma^M\,\Delta^M}_{\alert{\mathsf{concat}}} && (\sigma^M ;^M \tau^M) ,_{\Sub}^M \alert{\transfib{}{([;]^M_{\Ty})^{-1}}{\color{black} \subM{t^M}{\sigma^M}}} \\
  \pi_1\beta^M              & : \pi_1^M (\sigma^M ,^M_{\Sub} t^M)    && =^{\Sub^M\,\Gamma^M\,\Delta^M}_{\alert{\pi_1\beta}}      && \sigma^M \\
  \pi_2\beta^M              & : \pi_2^M (\sigma^M ,^M_{\Sub} t^M)    && =^{\Tm^M\,\Gamma^M}_{\alert{[\pi_1,]^M_{\Ty}, \pi_2\beta}}       && t^M \\
  \cdot\eta^M               & : \sigma^M                             && =^{\Sub^M\,\Gamma^M\,\emptyctx^M_{\Ctx}}_{\alert{\cdot\eta}}  && \emptysub_{\Sub}^M \\
  \pi\eta^M                 & : \sigma^M                             && =^{\Sub^M\,\Gamma^M\,(\Delta^M ,^M_{\Ctx} A^M)}_{\alert{\pi\eta}} {} && \pi_1^M \sigma^M ,_{\Sub}^M \pi_2^M \sigma^M \\
  [\idS]^M                  & : t^M [ \idS ]^M                      && =^{\Tm^M\,\Gamma^M}_{\alert{[\idS]^M_{\Ty}, [\idS]}} && t^M \\
  [;]^M                     & : t^M [ \sigma^M ;^M \tau^M ]^M       && =^{\Tm^M\,\Gamma^M}_{\alert{[;]^M_{\Ty}, [;]}} && \subM{\subM{t^M}{\tau^M}}{\sigma^M}
\end{alignat*}

>>>>

The structural rules for type substitution are stipulated by following equality constructors:
\begin{alignat*}{9}
  [\idS]_{\Ty} &: \implicit{\Gamma, A} \; && \sub{A}{\idS} && =^{\Ty\,\Gamma} A
  \qquad\qquad
  [;]_{\Ty}  : \implicit{\Gamma, \Delta, \Theta, \sigma, \tau, A}\; \sub{A}{\sigma ; \tau} =^{\Ty\,\Gamma} \sub{\sub{A}{\tau}}{\sigma}
  \\
  []\UU &: \implicit{\Gamma, \Delta, \sigma} \; && \sub{\UU}{\sigma} && =^{\Ty\,\Gamma} \UU
\end{alignat*}
%
%\begin{alignat*}{3}
%  [\idS]_{\Ty} & : \implicit{\Gamma, A}                               && \sub{A}{\idS}           && =^{\Ty\,\Gamma}\;A \\
%  [;]_{\Ty}    & : \implicit{\Gamma, \Delta, \Theta, \sigma, \tau, A}\;&& \sub{A}{\sigma ; \tau}  && =^{\Ty\,\Gamma}\; \sub{(\sub{A}{\tau})}{\sigma} \\
%  []\UU      & : \implicit{\Gamma, \Delta, \sigma}                    && \sub{\UU}{\sigma}        && =^{\Ty\,\Gamma}\;\UU
%\end{alignat*}

@@

However, we still have a little bit more work to do before we can declare victory: as we are defining a function on a quotient inductive type, we have to prove that $\subTy{\blank}{\blank}$ is coherent, i.e., that it respects all equality constructors.
%\LT[noinline]{Should we mention any general condition for coherence?}
%\Fredrik[noinline]{Is there one?}
%\LT[noinline]{I think Cubical Agda has sort of conditions? (Cubical Agda does not support rewriting)}
\begin{proposition}\label{prop:coherence-1}
  For every equality constructor $\cc : \sigma = \tau$, we have $\subTy{A}{\sigma} = \subTy{A}{\tau}$.
  \danger
\end{proposition}
\begin{proof}
  The equality constructors $[\idS]_T$, $[;]_T$, and $[]\UU$ turn into definitional equalities for type substitution, so it remains to show that the following identities
  \begin{align*}
    \subTy{A}{\sigma;\idS}          & = \subTy{A}{\sigma},
                                    & \subTy{A}{\idS;\sigma}          & = \subTy{A}{\sigma},
                                     & \subTy{A}{(\sigma;\tau);\gamma} & = \subTy{A}{\sigma; (\tau; \gamma)}, \\
    \subTy{A}{\pi_1(\sigma, t)}      & = \subTy{A}{\sigma}, \\
    \subTy{A}{\sigma; (\tau, t)}    & = \subTy{A}{(\sigma;\tau), \sub{t}{\sigma}},
                                     & \subTy{A}{\sigma}               & = \subTy{A}{\emptyctx},
                                     & \subTy{A}{\sigma}               & = \subTy{A}{\pi_1\sigma, \pi_2\sigma}
  \end{align*}
  hold propositionally.
  The first two identities of the first row hold strictly by \eqref{eq:def-type-subst-1} and \eqref{eq:def-type-subst-2}, while both sides of the third reduce to $\subTy{\subTy{\subTy{A}{\gamma}}{\tau}}{\sigma}$ by \eqref{eq:def-type-subst-2}.
  %
  The second row is clause \eqref{eq:def-type-subst-3} of the definition of $\subTy{\blank}{\blank}$.
  %
  We show the remaining cases by induction on $A$.
  However, $\UU$ is the only constructor for $\Ty\;\Gamma$ and each of the identities with $A \judgeq \UU$ reduce to $\UU$.
\end{proof}
\begin{remark}\label{re:coherence-proof}
  In the above proof the first four identities hold strictly by \eqref{eq:def-type-subst-1}--\eqref{eq:def-type-subst-4}, while the last three follow from \eqref{eq:def-type-subst-5} by induction on $A$.
  As we add other type formers, the first four identities will always hold, but the last three will depend on how type substitution is defined.
\end{remark}

@@

\begin{minipage}{.6\textwidth}
\begin{alignat*}{5}
  %\Tel               & :                   && \Ctx \to \Set                                     \\
  \emptytel          & : \implicit{\Gamma}\; && \Tel\;\Gamma \\
  \blank,\blank      & : \implicit{\Gamma}\; && (\Xi : \Tel\;\Gamma) \to \Ty\;(\Gamma \dplus \Xi) \to \Tel\;\Gamma
\end{alignat*}
\end{minipage}
\begin{minipage}{.4\textwidth}
\begin{alignat*}{5}
  %\blank & \dplus\blank     && :       (\Gamma : \Ctx) (\Xi : \Tel\;\Gamma) \to \Ctx \\
  \Gamma & \dplus \emptytel && \reduce \Gamma \\
  \Gamma & \dplus (\Xi, A)  && \reduce (\Gamma \dplus \Xi) , A
\end{alignat*}
\end{minipage}
\vskip\belowdisplayskip\noindent
Equipped with telescopes, we can now generalise single lifting with type \eqref{eq:type-of-lifting} to multiple lifting with type
\[
\blank\upuparrows \blank : (\sigma : \Sub\;\Gamma\;\Delta) \to (\Xi : \Tel\;\Delta)
\to \Sub\;(\Gamma \dplus \subTel{\Xi}{\sigma})\;(\Delta \dplus \Xi)
\]
which asks for telescope substitution $\subTel{\blank}{\blank} :  \Tel\;\Delta \to \Sub\;\Gamma\;\Delta \to \Tel\;\Gamma$ defined mutually:

@@

\begin{minipage}{.5\textwidth}
\begin{alignat*}{3}
  \subTel{\Xi}{\idS}          & \reduce \Xi \\
  \subTel{\Xi}{\sigma ; \tau} & \reduce \subTel{\subTel{\Xi}{\tau}}{\sigma}
\end{alignat*}
\end{minipage}
\begin{minipage}{.5\textwidth}
\begin{alignat*}{5}
  & \idS          && \upuparrows \Xi && \reduce \idS \\
  & (\sigma;\tau) && \upuparrows \Xi && \reduce (\sigma \upuparrows \subTel{\Xi}{\tau}) ; (\tau \upuparrows \Xi)
\end{alignat*}
\end{minipage}\vskip\belowdisplayskip\noindent
%whose confluence property is easy to see.

@@

Finally, we note that for $t : \Tm\,\Gamma\,(\Pi\,A\,B)$ and $u : \Tm\,\Gamma\,A$, also `ordinary' application
$t \mathop{\$} u \defeq \subTm{(\mathsf{app}\,t)}{\idS , t}$ of type $\Tm\,\Gamma\,(\sub{B}{\idS , t})$ can be defined without any transport. This is in contrast to the QIIT situation, where ordinary application is defined as $t \mathop{\$} u \defeq \sub{(\mathsf{app}\,t)}{\idS, \alert{\transfib{\Tm\;\Gamma}{[\idS]_{\Ty}^{-1}}{\color{black}t}}}$.
%
%\begin{alignat*}{3}
%  \left<\blank\right> & : \implicit{\Gamma, A} \Tm\;\Gamma\;A \to \Sub\;\Gamma\;(\Gamma, A) \\
%  \left< t \right>    & \defeq (\idS , t),
%\end{alignat*}
%used to define the ordinal application: $t \mathop{\$} u \defeq [ \left< u \right> ]_{\Tm} (\mathsf{app}\,t)$, 
%can be introduced without any transport, which is $(\idS, \alert{\transfib{\Tm\;\Gamma}{[id]_T^{-1}}{\color{black}t}})$ in its QII definition.
Similarly, the structural rule for substitution on $\mathsf{app}$
\[
  \sub{(\mathsf{app}\;t)}{\sigma \uparrow A} \overset{\Pi\beta^{-1}}{=}
  \mathsf{app}\,\lambda(\sub{(\mathsf{app}\;t)}{\sigma \uparrow A}) \overset{[]\lambda^{-1}}{=}
  \mathsf{app}\,\left(\sub{(\lambda\,(\mathsf{app}\,t))}{\sigma}\right) \overset{\Pi\eta^{-1}}{=}
  \mathsf{app}\,(\sub{t}{\sigma})
\]
%\begin{align*}
%  \sub{(\mathsf{app}\;t)}{\sigma \uparrow A} & = \mathsf{app}\,\lambda(\sub{(\mathsf{app}\;t)}{\sigma \uparrow A}) && \text{by $\Pi\beta^{-1}$} \\
%                                              & = \mathsf{app}\,\left(\sub{(\lambda\,(\mathsf{app}\,t))}{\sigma}\right) && \text{by $[]\lambda^{-1}$} \\
%                                              & = \mathsf{app}\,(\sub{t}{\sigma}) && \text{by $\Pi\eta^{-1}$.}
%\end{align*}
can be derived without transport, since \eqref{eq:def-type-subst-8} is now strict.

@@

\begin{remark}
  To state the methods, we cannot avoid using transports in identities that would need transports in its QIIT definition of a type theory --- for example, we have
  \[
    [\idS]_{\Tm}^M : \implicit{\Gamma, \Gamma^M, t, t^M} \;\subTmM{t^M}{\idS^M} =^{\Tm^M\,\Gamma^M\,\blank\,t}_{\alert{[\idS]^M_{\Ty}}} \; t^M
  \]
  since identities occurring in a method %(e.g.\ $[\idS]_{\Ty}^M$)
  are propositional just like equality constructors in its QIIT definition.
  However, in the elimination principle for the QIIT definition, the situation is even worse, as we in general need yet another layer of transports for the methods to typecheck over the QIIT constructors.

@@

The elimination principle computes by recursively interpreting each constructor using the corresponding method. For example, we have
\begin{alignat*}{5}
%  \elim_{\Ctx}\;\emptyctx   & \reduce \emptyctx_{\Ctx}                                    && : \Ctx^M\;\emptyctx  \\
  \elim_{\Ctx}\;(\Gamma, A) & \reduce (\elim_{\Ctx}\,\Gamma),_{\Ctx}^{M} (\elim_{\Ty}\,A) . && %\\
%  \elim_{\Ty}\;\UU          & \reduce \UU^M && :\Ty^M\;\Gamma^M\;\UU \\
%  \elim_{\Sub}\;\emptysub   & \reduce \emptysub^M_{\Sub} &&: \Sub^M\;\Gamma^M\;\emptyctx^M_{\Ctx}\;\emptyctx \\
%                            & \vdots
\end{alignat*}
%
Interestingly, a key difference compared to its QIIT counterpart is that $\subTy{\blank}{\blank}$, $\subTm{\blank}{\blank}$, and $\blank\uparrow\blank$ are functions rather than (formed by) constructors, meaning they cannot be eliminated inductively but instead become propositional identities.
%type substitution $[\blank]\blank$ was a constructor and $[\sigma]\,A$ was eliminated to $[\elim_{\Sub}\,\sigma]^M_{\Ty}\,(\elim_{\Ty}\,A)$; similarly for term substitution.
\begin{restatable}{proposition}{propProvableFunc}\label{prop:provable-function-clauses}
  Given the motives and the methods, we can prove
\begin{alignat*}{5}
  \elim_{\Ty}[] & : \elim_{\Ty}\,(\subTy{A}{\sigma}) && = && \subTyM{(\elim_{\Ty}\,A)}{\elim_{\Sub}\,\sigma} \\
  \elim_{\Tm}[] & : \elim_{\Tm}\,(\subTm{t}{\sigma}) && =^{\Tm^M\,\Gamma^M\,\blank\,t}_{\alert{\elim_{\Ty}[]}} {} && \subTmM{(\elim_{\Tm}\,t)}{\elim_{\Sub}\,\sigma} \\
  \elim_{\Sub}[] & : \elim_{\Sub}(\sigma\uparrow A) && =^{\Sub^M\,(\Gamma^M,^M \blank)\,(\Delta^M, A^M)\,(\sigma \uparrow A)}_{\alert{\elim_{\Ty}[]}} {} && (\elim_{\Sub} \sigma) \uparrow^M (\elim_{\Ty}\,A)
\end{alignat*}
\end{restatable}

@@

\begin{alignat*}{5}
  ;\text{-}\textsf{idr}^M   & : \sigma^M ;^M \idS^M && =^{\Sub^M\,\Gamma^M\,\Delta^M}_{\alert{;\text{-}\mathsf{idr}}} && \sigma^M \\
  ;\text{-}\textsf{idl}^M   & : \idS^M ;^M \sigma^M && =^{\Sub^M\,\Gamma^M\,\Delta^M}_{\alert{;\text{-}\mathsf{idl}}} && \sigma^M \\
  ;\text{-}\textsf{assoc}^M & : (\sigma^M ;^M \tau^M) ;^M \gamma^M   && =^{\Sub^M\,\Gamma^M\,\Delta^M}_{\alert{;\text{-}\mathsf{assoc}}} && \sigma^M ;^M (\tau^M ;^M \gamma^M) \\
  \mathsf{concat}^M         & : \sigma^M ;^M (\tau^M ,_{\Sub}^M t^M) && =^{\Sub^M\,\Gamma^M\,\Delta^M}_{\alert{\mathsf{concat}}} && (\sigma^M ;^M \tau^M) ,_{\Sub}^M \alert{\transfib{}{([;]^M_{\Ty})^{-1}}{\color{black} \subM{t^M}{\sigma^M}}} \\
  \pi_1\beta^M              & : \pi_1^M (\sigma^M ,^M_{\Sub} t^M)    && =^{\Sub^M\,\Gamma^M\,\Delta^M}_{\alert{\pi_1\beta}}      && \sigma^M \\
  \pi_2\beta^M              & : \pi_2^M (\sigma^M ,^M_{\Sub} t^M)    && =^{\Tm^M\,\Gamma^M}_{\alert{[\pi_1,]^M_{\Ty}, \pi_2\beta}}       && t^M \\
  \cdot\eta^M               & : \sigma^M                             && =^{\Sub^M\,\Gamma^M\,\emptyctx^M_{\Ctx}}_{\alert{\cdot\eta}}  && \emptysub_{\Sub}^M \\
  \pi\eta^M                 & : \sigma^M                             && =^{\Sub^M\,\Gamma^M\,(\Delta^M ,^M_{\Ctx} A^M)}_{\alert{\pi\eta}} {} && \pi_1^M \sigma^M ,_{\Sub}^M \pi_2^M \sigma^M \\
  [\idS]^M                  & : \subM{t^M}{\idS}                     && =^{\Tm^M\,\Gamma^M}_{\alert{[\idS]^M_{\Ty}, [\idS]}} && t^M \\
  [;]^M                     & : \subM{t^M}{\sigma^M ;^M \tau^M}      && =^{\Tm^M\,\Gamma^M}_{\alert{[;]^M_{\Ty}, [;]}} && \subM{\subM{t^M}{\tau^M}}{\sigma^M}
\end{alignat*}
